---
title: "Project"
author: "Carlo Carbonilla"
date: "13/05/2021"
output: pdf_document
---
```{r}
library(dplyr)
```

Question 1.

(a)

```{r}
sales <- read.csv(file = 'STAT702_project_sales_data.csv')
head(sales)
```


Convert week column from character to date type
```{r}
sales$week <- as.Date(sales$week, format =  "%d/%m/%y")
head(sales)
```

Filter dataframe to dates in 2012 and with the product sku of 216419
```{r}
productSales <- sales %>% filter(sku_id == 216419)
productSales
```

```{r}

```


Question 2.

Filter dataframe to dates in 2012 and with the product sku of 216419
```{r}
productSales2012 <- sales %>% filter(week >= as.Date("2012-01-01"), week <= as.Date("2012-12-31"), sku_id == 216419)
head(productSales2012)
```

i.
```{r}
productSales2012
```

Setup cost = $130 per order
Holding cost = $1.50 per unit per year

k = order or setup cost
A = Number of items demanded per annum
c = purchase cost per item
h = holding cost per item per annum
Q* = optimum order quantity (number of items)
T = time between order = length of the inventory cycle = Q/A


Tc = Annual inventory cost = Annual order cost + Annual holding cost
   = 130*Q + 1.5
























